---
title: "RSA example"
output: html_document
---

We will see what different layers of the  model do. 
First we need to import all the functions and objects

Source code that creates objects and utterance vectors
```{r echo=FALSE}
## All Utterances
# All possible utterances (i.e. object features) that can be handled.
# Here, we assume a 3x3 matrix (three feature types with three expressions each)
allUtterances <- c('cloud', 'circle', 'square', 'solid', 'striped', 'dotted', 'blue', 'red', 'green')
allUtterancesNew1 <- c('cloud', 'circle', 'square', 'solid', 'striped', 'polka-dotted', 'blue', 'red', 'green')
allFeatureTypesNew1 <- c('shape','pattern','color')
allUttMatrix <- matrix(allUtterances, ncol=3, byrow=TRUE)
##
## All Objects
# all object matrix contains 3^3 types of objects.
# the matrix essentially specifies the 3 feature expressions for each object
# thus, the matrix maps objects to matching utterances
# all Objects implements the strings, 
# allObjectsToUtterancesMappings encodes the index mappings
allObjects <- matrix('',27,3)
allObjectsToUtterancesMappings <- matrix(0,27,3)
for(index in c(1:27)) {
  #  print(c(1+((index-1)%%3), 1+floor(((index-1)%%9)/3), 1+floor((index-1)/9)))
  allObjects[index,1] <- allUttMatrix[1,1+((index-1)%%3)]
  allObjects[index,2] <- allUttMatrix[2,1+floor(((index-1)%%9)/3)]
  allObjects[index,3] <- allUttMatrix[3,1+floor((index-1)/9)]
  allObjectsToUtterancesMappings[index,1] <- 1+((index-1)%%3)
  allObjectsToUtterancesMappings[index,2] <- 4+floor(((index-1)%%9)/3)
  allObjectsToUtterancesMappings[index,3] <- 7+floor((index-1)/9)
}

## 
## The relevant utterances are determined given currentObjects
# valid utterances correspond to all features present in the current objects!
determineValidUtterances <- function(currentObjects) {
  validUtterances <- c()
  for(i in c(1:length(currentObjects))) {
    validUtterances <- c(validUtterances, allObjectsToUtterancesMappings[currentObjects[i],])
  }
  validUtterances <- sort(unique(validUtterances))
  return(validUtterances)
}

###
## No preference is encoded with 4, whereas a specific feature expression preference is encode 
# by the respective index value
# get feature-respective priors returns general feature respective priors for all 3 features
# @deprecated (not used currently!)
getFeatureRespectivePriors <- function(softAddProb) {
  featureRespectivePriors <- list()
  for(i in c(1:3)) { ## for all three features generate a preference matrix
    m <- matrix(0,4,3)
    for(fPref in c(1:3)) {
      m[fPref,fPref] <- 1
      m[fPref,] <- m[fPref,] + softAddProb
      m[fPref,] <- m[fPref,] / sum(m[fPref,])
    }
    m[4,] <- 1/3
    featureRespectivePriors[[i]] <- m
  }
  return(featureRespectivePriors)
}

##
## Determining the specifc mapping of objects to utterances that applies given currentObjects
# mapping current objects to utterances
determineObjectToUtterancesMapping <- function(currentObjects) {
  mapObjToUtt <- matrix(0, length(currentObjects), 3)
  for(i in c(1:length(currentObjects))) {
    mapObjToUtt[i,] <- allObjectsToUtterancesMappings[currentObjects[i],]
  }
  return(mapObjToUtt)
}

##
# Determining the corresponding mappings from all relevant utterances to objects
# parameter notObeyInst determines if the instruction does not need to be obeyed (0=full obedience: -> infty  =full instruction ignorance) 
determineUtteranceToObjectProbabilities <- function(consideredUtterances, currentObjects, 
                                                    mapObjToUtt, notObeyInst) {
  mapUttToObj <- list()
  mapUttToObjProbs <- matrix(notObeyInst, length(consideredUtterances), length(currentObjects))
  for(utt in rep(1:length(consideredUtterances)) ) {
    # determine array of all objects that match the utterance:
    #  take utterance index utt, 
    #  return the indices of those rows of mapObjToUtt (=indices of objects) that have the feature denoted by utt
    mapUttToObj[[utt]] = ((which(mapObjToUtt[,] == consideredUtterances[utt])-1)%%nrow(mapObjToUtt))+1
    #a (hopefully) equivalent alternative to the previous line: 
    #  mapUttToObj[[utt]] = which(apply(mapObjToUtt, 1, function(currRow) (utt %in% currRow)))
  
  # Alternative code for the loop below
    
  #  for(obj in mapUttToObj[[utt]]) {
  #    mapUttToObjProbs[utt,obj] <- mapUttToObjProbs[utt,obj] + 1;
  #  }
    
    for(i in rep(1:length(mapUttToObj[[utt]]))) {
      mapUttToObjProbs[utt,mapUttToObj[[utt]][i]] <- mapUttToObjProbs[utt,mapUttToObj[[utt]][i]] + 1;
     }
     
    
    
    mapUttToObjProbs[utt,] <- mapUttToObjProbs[utt,] / sum(mapUttToObjProbs[utt,])# length(mapUttToObj[[utt]])
  }
  return(mapUttToObjProbs)
}

##
## Priors on object preferences - automatically derived from considered utterances
#    (i.e. derived from all relevant features)
# type == 0: hard priors; type > 0: soft prior with specified softness
# returns a list of preference priors for all considered features, i.e. utterances, 
# as well as for "no preference" whatsoever, i.e., uniform prior over all three objects
getObjectPreferencePriors <- function(consideredUtterances, currentObjects, type, mapUttToObjProbs) {
  objectPreferenceHardPriors <- list()
  for(utt in rep(1:length(consideredUtterances)) ) {
    objectPreferenceHardPriors[[utt]] <- mapUttToObjProbs[utt,]
  }
  objectPreferenceHardPriors[[length(consideredUtterances)+1]] = # Adding an extra row with flat prior over objects
    rep(1/length(currentObjects), length(currentObjects) )
  # soft preferences with uniform choice fusion. 
  softAddProb <- type
  objectPreferenceSoftPriors <- list()
  for(utt in rep(1:(length(consideredUtterances)+1)) ) {
    objectPreferenceSoftPriors[[utt]] <- objectPreferenceHardPriors[[utt]] + softAddProb
    objectPreferenceSoftPriors[[utt]] <- objectPreferenceSoftPriors[[utt]] / sum(objectPreferenceSoftPriors[[utt]])
  }
  return(objectPreferenceSoftPriors)
}



######## From Ella's Code:
# gives a full random matrix of all feature value preferences given a certain ratio i.e. c(0, 1/3, 2/3)
getAllUtterancePref <- function(ratio) {
  allUttNum <- c(1:9)
  allUtterancePref <- matrix(allUttNum, length(allUttNum), 3)
  sampledRatio <- c(sample(ratio), sample(ratio), sample(ratio))
  repRatio <- c(ratio, ratio, ratio)
  for(utt in rep(1:length(allUtterances))){
    allUtterancePref[utt,1] <- utt
    allUtterancePref[utt,2] <- allUtterances[utt]
    # allUtterancePref[utt,3] <- as.numeric(sampledRatio[utt])
    allUtterancePref[utt,3] <- as.numeric(repRatio[utt])
  }
  return(allUtterancePref)
}

determineAllFeaValues <- function(currentObjects) {
  allFeaUtterances <- c()
  for(i in c(1:length(currentObjects))) {
    allFeaUtterances <- c(allFeaUtterances, allObjectsToUtterancesMappings[currentObjects[i],])
  }
  return(allFeaUtterances)
}
# get a matrix of all preferences for the relevantUtterances
getMapUttToPref <- function(relevantUtterances, allObjects, allUtterancePref){
  mapUttToPref <- allUtterancePref
  allUtteranceNum <- c(1:9)
  notRelevantUtt <- setdiff(allUtteranceNum, relevantUtterances)
  for (rowNum in rep(1:length(notRelevantUtt))){
    mapUttToPref <- mapUttToPref[-which(mapUttToPref[,1] == notRelevantUtt[rowNum]),]
  }
  return (mapUttToPref)
}

getObjectPreferencePriorsWithUttToPref <- function(consideredUtterances, currentObjects, type, mapUttToObjProbs, mapUttToPref) {
  objectPreferenceHardPriors <- list()
  for(utt in rep(1:length(consideredUtterances)) ) {
    objectPreferenceHardPriors[[utt]] <- mapUttToObjProbs[utt,]#*as.numeric(mapUttToPref[utt,3])
  }
  #nopreference case
  objectPreferenceHardPriors[[length(consideredUtterances)+1]] = rep(1/length(currentObjects), length(currentObjects) )
  
  #  soft preferences with uniform choice fusion. 
  softAddProb <- type
  objectPreferenceSoftPriors <- list()
  for(utt in rep(1:(length(consideredUtterances)+1)) ) {
    objectPreferenceSoftPriors[[utt]] <- objectPreferenceHardPriors[[utt]] + softAddProb
    objectPreferenceSoftPriors[[utt]] <- objectPreferenceSoftPriors[[utt]] / sum(objectPreferenceSoftPriors[[utt]])
  }
  return(objectPreferenceSoftPriors)
}
# creates a matrix containing all preferences for the object choice of the listener, 
# depending on the present objects and the target feature and the preferences
# of the listener
# targetfeature has to be numeric
# +1 row without preferences
getMapUttToObjToPref <- function(currentObjects, targetFeature, relevantUtterances, allUtterancePref, allObjects, mapUttToPref){
  isUnique <- matrix(FALSE, nrow = length(relevantUtterances)+1, ncol = length(currentObjects))
  mapUttToObjToPref <- matrix(0, nrow = length(relevantUtterances)+1, ncol = length(currentObjects))
  objectSpecific <- matrix("", nrow= 3, ncol = 3)
  for(obj in rep(1:3)){
    objectSpecific[obj,] <- allObjects[currentObjects[obj],]
  }
  countedUttObj <- table(objectSpecific)
  relevantUttWords <- relevantUtterances
  for(utt in rep(1:length(relevantUttWords))){
    relevantUttWords[utt] <- mapUttToPref[utt,2]
    index <- which(objectSpecific == relevantUttWords[utt], arr.ind=TRUE)[,"row"]
    for(ind in rep(1:length(index))){
      isUnique[utt,index[ind]] <- TRUE
    }
  }
  # cat(print(relevantUttWords))
  # cat(print(isUnique))
  for(row in rep(1:length(relevantUtterances))){
    for(col in rep(1:length(currentObjects))){
      if(isUnique[row,col]){
        targetFeatureValue <- objectSpecific[col,targetFeature]
        targetFeatureValuePref <- mapUttToPref[which(mapUttToPref[,2]==targetFeatureValue),3]
        if(length(targetFeatureValuePref)==0) {
          print(c("length is zero!?",targetFeatureValue,targetFeatureValuePref))
          print(allUtterances[relevantUtterances])
          print(allObjects[currentObjects,])
        }
        mapUttToObjToPref[row,col] <- targetFeatureValuePref
      }
    }
  }
  # cat(print(isUnique))
  mapUttToObjToPref[length(relevantUtterances)+1,]<- 0.33
  if(targetFeature == "shape" || targetFeature == 1){
    notRelevantUtt <- c(1, 2, 3)
  } else  if(targetFeature == "pattern" || targetFeature == 2){
    notRelevantUtt <- c(4, 5, 6)
  } else  if(targetFeature == "color" || targetFeature == 3){
    notRelevantUtt <- c(7, 8, 9)
  }
  currentNotRelevantUtt <- which(as.numeric(mapUttToPref[,1]) %in% notRelevantUtt)
  for (row in rep(1:length(currentNotRelevantUtt))){
    mapUttToObjToPref[currentNotRelevantUtt[row],] <- 0
  }
  return (mapUttToObjToPref)
}


```

```{r}
allObjects
```

```{r}
allUtterances
```

```{r}
allFeatureTypesNew1
```

Let's create a trial with 3 objects, and define what was the utterance, and what properties we'd like to learn about

```{r}
currentObjects <- c(1,2,6)
allObjects[currentObjects,]
utterance <- 4
targetFeature <- 1
obj <- 1
```

## Simple listener

```{r}
simpleListener <-
  function(utterance,
           mapUttToObjProbs,
           listenerObjectPreferences) {
    objPosterior <-
      mapUttToObjProbs[utterance, ] * (listenerObjectPreferences + 1e-100)
    if (sum(objPosterior) == 0) {
      return(objPosterior)
    }
    return(objPosterior / sum(objPosterior))
  }
```

Now we take the current objects and see what the literal listener does. First we need to find all the variables. 

```{r}
determineObjectToUtterancesMapping <- function(currentObjects) {
  mapObjToUtt <- matrix(0, length(currentObjects), 3)
  for(i in c(1:length(currentObjects))) {
    mapObjToUtt[i,] <- allObjectsToUtterancesMappings[currentObjects[i],]
  }
  return(mapObjToUtt)
}
```

```{r}
mapObjToUtt <- determineObjectToUtterancesMapping(currentObjects)
```

```{r}
determineUtteranceToObjectProbabilities <- function(consideredUtterances, currentObjects, 
                                                    mapObjToUtt, notObeyInst) {
  mapUttToObj <- list()
  mapUttToObjProbs <- matrix(notObeyInst, length(consideredUtterances), length(currentObjects))
  for(utt in rep(1:length(consideredUtterances)) ) {
    mapUttToObj[[utt]] = ((which(mapObjToUtt[,] == consideredUtterances[utt])-1)%%nrow(mapObjToUtt))+1
    for(i in rep(1:length(mapUttToObj[[utt]]))) {
      mapUttToObjProbs[utt,mapUttToObj[[utt]][i]] <- mapUttToObjProbs[utt,mapUttToObj[[utt]][i]] + 1;
     }
    mapUttToObjProbs[utt,] <- mapUttToObjProbs[utt,] / sum(mapUttToObjProbs[utt,]) 
  }
  return(mapUttToObjProbs)
}
```

We need to find out what feature values are present

```{r}
determineValidUtterances <- function(currentObjects) {
  validUtterances <- c()
  for(i in c(1:length(currentObjects))) {
    validUtterances <- c(validUtterances, allObjectsToUtterancesMappings[currentObjects[i],])
  }
  validUtterances <- sort(unique(validUtterances))
  return(validUtterances)
}
```

```{r}
validUtterances <- determineValidUtterances(currentObjects)
```

Now we have everything to determine the probabilities that a literal listener will pick an object if a certain utterance is spoken

```{r}
mapUttToObjProbs <- determineUtteranceToObjectProbabilities(validUtterances, currentObjects, mapObjToUtt, 0)
```

Still missing one piece to calculate the listener predictions

```{r}
getObjectPreferencePriors <- function(consideredUtterances, currentObjects, type, mapUttToObjProbs) {
  objectPreferenceHardPriors <- list()
  for(utt in rep(1:length(consideredUtterances)) ) {
    objectPreferenceHardPriors[[utt]] <- mapUttToObjProbs[utt,]
  }
  objectPreferenceHardPriors[[length(consideredUtterances)+1]] = # Adding an extra row with flat prior over objects
    rep(1/length(currentObjects), length(currentObjects) )
  # soft preferences with uniform choice fusion. 
  softAddProb <- type
  objectPreferenceSoftPriors <- list()
  for(utt in rep(1:(length(consideredUtterances)+1)) ) {
    objectPreferenceSoftPriors[[utt]] <- objectPreferenceHardPriors[[utt]] + softAddProb
    objectPreferenceSoftPriors[[utt]] <- objectPreferenceSoftPriors[[utt]] / sum(objectPreferenceSoftPriors[[utt]])
  }
  return(objectPreferenceSoftPriors)
}
```

```{r}
objectPreferenceSoftPriors <- getObjectPreferencePriors(validUtterances, currentObjects, 0, mapUttToObjProbs)
```

Now finally the literal listener!

```{r}
simpleListener(utterance, mapUttToObjProbs, objectPreferenceSoftPriors[[utterance]])
```


## Pragmatic speaker

```{r}
simplePragmaticSpeaker <-
  function(utterance,
           obj,
           preferencesPrior,
           relevantUtterances,
           currentObjects,
           mapUttToObjProbs,
           objectPreferenceSoftPriors) {
    prefPost <- rep(0, length(preferencesPrior)) 
    for (pref in c(1:length(preferencesPrior))) {
      if (preferencesPrior[pref] > 0) {
        pp <-
          simpleListener(utterance,
                         mapUttToObjProbs,
                         objectPreferenceSoftPriors[[pref]])
        prefPost[pref] <- pp[obj] * preferencesPrior[pref]
      }
    }
    if (sum(prefPost) == 0) {
      # no evidence for any preferences... -> no inference
      return(preferencesPrior)
    }
    return(prefPost / sum(prefPost))
  }
```

We need to first calculate the preferences prior

```{r}
getPreferencesPrior <- function(targetFeature) {
  preferencesPrior <- c(rep(0, 9))
  index <- targetFeature * 3
  indices <- c(index-2, index - 1, index)
  preferencesPrior[indices] <- 1
  return(preferencesPrior / sum(preferencesPrior))
}
```

```{r}
preferencesPrior <- getPreferencesPrior(targetFeature)
```

```{r}
speaker <-simplePragmaticSpeaker(utterance, obj, preferencesPrior, validUtterances, currentObjects, mapUttToObjProbs, objectPreferenceSoftPriors)
round(speaker, 2)
```

Please work out why we get a probability of 0.33 for the third object.

## Picking the most useful utterance

```{r}
simpleBestInfGainUtterance <-
  function(preferencesPrior,
           relevantUtterances,
           currentObjects,
           mapUttToObjProbs,
           objectPreferenceSoftPriors,
           klValueFactor = 1,
           utterancePrior) { # added argument for illustration
    postUttGPrefPrior <- rep(0, length(relevantUtterances))
    for (utt in c(1:length(relevantUtterances))) {
      prefPostAll <- rep(0, length(preferencesPrior))
      for (pref in c(1:length(preferencesPrior))) {
        prefPost <- 0
        for (obj in c(1:length(currentObjects))) {
          if (mapUttToObjProbs[utt, obj] > 0) {
            if (preferencesPrior[pref] > 0) {
              featurePrefsPosterior <-
                simplePragmaticSpeaker(
                  utt,
                  obj,
                  preferencesPrior,
                  relevantUtterances,
                  currentObjects,
                  mapUttToObjProbs,
                  objectPreferenceSoftPriors
                )
              KLvalue <-
                KLdivergence(preferencesPrior, featurePrefsPosterior)
              prefPost <- prefPost +  mapUttToObjProbs[utt, obj] *
                objectPreferenceSoftPriors[[pref]][obj] *
                utterancePrior[utt] *  preferencesPrior[pref] *
                exp(klValueFactor * KLvalue)
            }
          }
        }
        if (prefPost > 0) {
          prefPostAll[pref] <- prefPost
        }
      }
      postUttGPrefPrior[utt] <- sum(prefPostAll)
    }
    if (sum(postUttGPrefPrior) == 0)
      # no gain from any utterance...
      return(rep(1 / length(relevantUtterances), length(relevantUtterances)))
    return(postUttGPrefPrior / sum(postUttGPrefPrior))
  }
```

Before we can run it we need to define utterance prior

```{r}
relevantUtterances <- validUtterances
 utterancePrior <- rep(0,length(relevantUtterances))
  irrelevantIndices <- which(relevantUtterances>(3*(targetFeature-1)) & relevantUtterances<(3*targetFeature + 1))
  validUtterances <- relevantUtterances[-irrelevantIndices]
  utterancePriorShort <- rep (1/length(validUtterances),length(validUtterances)) 
  utterancePrior[-irrelevantIndices] <- utterancePriorShort
```

And the Kullback-Leibler divergence function

```{r}

KLdivergence <-
  function(p, q) {
    toleranceOffset <- 1e-20
    return(max(0, sum(p * (log(
      (toleranceOffset + p) / (toleranceOffset + q)
    )))))
  }
```

```{r}
bestUtterance <- simpleBestInfGainUtterance(preferencesPrior, relevantUtterances, currentObjects, mapUttToObjProbs, objectPreferenceSoftPriors, 1, utterancePrior)
round(bestUtterance,2)
```

